---
title: "Assignment_2 Lab1"
output: html_notebook
---

Assignment 2 Lab1 by Madhvi Jha

Applied Exercise 1
This exercise relates to the College data set, which can be found in the file College.csv. It contains a number of variables for 777 different universities and colleges in the US. The variables are

Private: Public/private indicator</br>
Apps: Number of applications received</br>
Accept: Number of applicants accepted</br>
Enroll: Number of new students enrolled</br>
Top10perc: New students from top 10% of high school class</br>
Top25perc: New students from top 25% of high school class</br>
F.Undergrad: Number of full-time undergraduates</br>
P.Undergrad: Number of part-time undergraduates</br>
Outstate: Out-of-state tuition</br
Room.Board: Room and board costs</br>
Books: Estimated book costs</br>
Personal: Estimated personal spending</br>
PhD: Percent of faculty with Ph.D.'s</br>
Terminal: Percent of faculty with terminal degree</br>
S.F.Ratio: Student/faculty ratio</br>
perc.alumni: Percent of alumni who donate</br>
Expend: Instructional expenditure per student<br>
Grad.Rate: Graduation rate</br>
Before reading the data into R, it can be viewed in Excel or a text editor.

**Part 1**</br>
Use the read.csv() function to read the data into R.</br> Call the loaded data college. Make sure that you have the directory set to the correct location for the data.
```{r}
# install.packages("ISLR")
library(ISLR)
library(MASS)
# college = read.csv("file name")
# Instead of running above command we are getting from package
data("College")
college = na.omit(College)
```

**Part 2**</br>
Look at the data using the fix() function. You should notice that the first column is just the name of each university. We don't really want R to treat this as data. However, it may be handy to have these names for later. Try the following commands:

**It is not needed since we are not loading from csv**
However, we could have run below code if done from csv
</br>
rownames(college) = college[, 1]</br>
fix(college)
</br>
```{r}
head(college)
```

**Part 3.1**</br>
Use the summary() function to produce a numerical summary of the variables in the data set.
```{r}
summary(college)

```
**Part 3.2**</br>
Use the pairs() function to produce a scatterplot matrix of the first ten columns or variables of the data. Recall that you can reference the first ten columns of a matrix A using A[, 1:10].
```{r}
pairs(college[,1:10])
```

